### 2024-03-12 paper/paper.qmd
paper so far

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Talia Fabregas
  - Fatimah Yunusa
  - Aamishi Sundeep Avarsekar
thanks: "Code and data are available at: https://github.com/taliafabs/US-Election-Forecast-2024.git"
date: today
date-format: long
abstract: "The 2024 U.S. Presidential Election ..."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(rstanarm)
library(readr)
library(ggplot2)
library(knitr)
library(arrow)
```

# Introduction

The 2024 U.S. Presidential Election will take place on Tuesday November 5 2024. Incumbent President Joseph R. Biden Jr. will seek a second term. Former President Donald J. Trump GOP nominee.

Delegates to secure the nomination, Nikki Haley has dropped out of GOPYou can and should cross-reference sections and sub-sections. We use @citeR and

The remainder of this paper is structured as follows. @sec-data discusses the survey and post-stratification data used.

Clear gap that needs to be filled ... what is the research gap and why is this important?

-   2024 US Presidential Election

-   how consequential this election is

-   what is on the ballot

-   women's rights

-   trump's project 2025 thing

If someone's done it before not on this data set not in this context

Previous groups have looked at ...

@citeR was usd

# Data {#sec-data}

```{r}
library(arrow)
library(dplyr)
survey_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/survey_analysis_data.parquet")
# poststrat_analysis_data <- read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
print(survey_analysis_data)

proportion_biden <- survey_analysis_data %>%
  filter(vote_biden == 1) %>%
  summarise(proportion = n() / nrow(survey_analysis_data))

proportion_biden_by_state <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(proportion_biden = mean(vote_biden == 1, na.rm = TRUE))

proportion_by_state <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(proportion = n() / nrow(survey_analysis_data))

proportion_biden_by_state
proportion_by_state

# only includes 7 observations from hawaii
# Hawaii is historically a Democratic stronghold
# only 4 samples from kansas
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-survey_presidential_preferences
#| fig-cap: The Distribution of Presidential Preferences, by gender and highest level of education
survey_analysis_data |>
  ggplot(aes(x = education_level, fill = vote24)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    x = "Highest education",
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

## Survey Data {#sec-survey}

Our survey data is from the America's Political Pulse Survey conducted by the Polarization Research Lab. The Polarization Research Lab is a research group founded by top political science research scholars at Dartmouth College, Stanford University, and the University of Pennsylvania, dedicated to applying scientific research methods to the study of democracy and political polarization @americaspoliticalpulse. The Polarization Research Lab studies political polarization across the United States by conducting the Political Pulse Survey each week and making the results available to anyone, in real time, via their website. The America's Political Pulse Survey interviews 1000 American adults each week and asks them questions that aim to track affective polarization, support for violations of democratic norms, and support for the use of political violence in America @americaspoliticalpulse. The Polarization Research Lab makes high-level results of their survey easily accessible and visible via their website's dashboard.

The specific survey data set used is the America's Political Pulse Survey Data Week 3 2024 January 12-19. It contains the responses of the 1000 American adults interviewed between January 12-19 2024. Demographic information about respondents, including sex, age, race, home state, employment status, marital status, ... is included. Participants were als asked who they voted for in the 2016 and 2020 U.S. presidential characteristics, what their party affiliation is, and what their political ideology is.

However, this data set has a few major limitations. It contains very few respondents from Hawaii, Wyoming, and Kansas. It contains zero respondents from Vermont.

The survey does not contain a question about preferred 2024 Presidential Candidate. In fact, it was conducted in January 2024, before Super Tuesday took place and Donald Trump had secured almost enough delegates to win the Republican nomination. At the time when this survey was conducted, former President Donald Trump, former South Carolina Governor Nikki Haley, and Florida Governor Ron DeSantis were all GOP Presidential primary candidates. Media personality Vivek Ramaswamy ended his primary campaign on January 15th, however this was in no way a clear indication that Trump had secured the nomination. As of March 11, 2024, former President Trump has secured 1078 out of the 1215 delegates needed to win the GOP nomination.

### Survey Data figures and tables go here

In order to gain a better understanding of our survey data, we conducted exploratory data analysis.

Within our survey data, the vote distribution between Trump and Biden varies by sex and race. As seen in @fig-survey_presidential_preferences1, Black females were more likely to vote for President Joe Biden (D-Delaware) than former President Donald Trump (R-Florida) in 2020, whereas white males were more likely to vote for former President Trump. This indicates that race might be a predictor of who an individual's preferred presidential candidate is.

```{r}
#| label: fig-survey_presidential_preferences1
#| fig-cap: The Distribution of Presidential Preferences, by gender and race
#| echo: false

survey_analysis_data |>
  ggplot(aes(x = races, fill = vote24)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    x = "Race",
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

```{r}
#| label: fig-electoral-map-survey
#| fig-cap: Electoral map based solely on the survey data
#| echo: false
library(ggplot2)
library(statebins)
survey_electoral_map <- survey_analysis_data %>% 
  group_by(state, vote_biden) %>% 
  summarise(n = n()) %>% 
  group_by(state) %>% 
  mutate(prop = n/sum(n)) %>% 
  filter(vote_biden == 1)

survey_electoral_map %>%
  mutate(statename = str_to_title(state)) %>%
  ggplot(aes(fill = prop, state = statename)) +
  geom_statebins() +
  scale_fill_gradient2(low = "#d12531", high = "#244999",
                       mid = "white", midpoint = mean(survey_electoral_map$prop)) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(fill = "Proportion Voting \nfor Biden",
       title = "Map of the US States that plan to support Joe Biden or Donald Trump")

# note extremely small sample sizes from Hawaii, Kansas, and Wyoming
# note that survey sample size was only 1000
```

## Post Stratification Data {#sec-poststrat}

@citeIPUMS

```{r}
#| label: fig-temp1
#| fig-cap: abc
#| echo: false

```

Talk more about it.

```{r}
#| label: fig-temp2
#| fig-cap: temp
#| echo: false
#| warning: false
#| message: false

# analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
# 
# analysis_data |> 
#   ggplot(aes(x = width, y = length)) +
#   geom_point(alpha = 0.8) +
#   theme_minimal() +
#   labs(x = "Wing width (mm)",
#        y = "Wing length (mm)")
```

Talk way more about it.

# Model

logistic regression

binary

predicts support for trump or biden

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

```{r survey_model, echo=F, message=F, fig.cap="Results of Regression Model"}
library(rstanarm)
library(arrow)
library(dplyr)

# poststrat_analy <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
# 
# poststrat_data_cells <-
#   poststrat_analysis_data |>
#   count(state, sex, age_group, education_level)
# 
# poststrat_data_cells <-
#   poststrat_data_cells |>
#   mutate(prop = n / sum(n),
#          .by = stateicp)
# 
# poststrat_data_cells

```

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

The logistic regression used to predict ... is as follows:

```{=tex}
\begin{align}
log(\frac{p}{1-p}) = \beta_0 + \beta_1 \times x_{sex} + \beta_2 \times x_{age\_bracket} +  \beta_3 \times x_{races} + \\
\beta_3 \times x_{education\_level} + \beta_4 \times x_{state} + \beta_5 
\times x_{urban} + \beta_5 \times x_{race\_black}
\end{align}
```
We run our model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

We use the `predict()` function in R @citeR to apply our logistic regression model to the ACS Census data @citeIPUMS.

Model weaknesses: logistic regression output is binary so it does not include the options of not voting or voting for a third-party candidate.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

presidential_vote_model <-
  readRDS(file = here::here("models/presidential_vote_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of presidential vote based on ..."
#| warning: false

modelsummary::modelsummary(
  list(
    "Presidential Vote Model" = presidential_vote_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## Popular Vote Projection {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Electoral College Projection

## Polarization and America's Urban-Rural Divide

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Correlation Map

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

# pp_check(first_model) +
#   theme_classic() +
#   theme(legend.position = "bottom")
# 
# posterior_vs_prior(first_model) +
#   theme_minimal() +
#   scale_color_brewer(palette = "Set1") +
#   theme(legend.position = "bottom") +
#   coord_flip()
```

Model plot (still need to fix this)

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelplot1
#| tbl-cap: "Model Plot"
#| warning: false

# modelplot(presidential_vote_model, conf_level = 0.9) +
#   labs(x = "90 per cent credibility interval")

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2
# 
# plot(first_model, "trace")
# 
# plot(first_model, "rhat")
```

\newpage

# References