### 2024-03-15 paper/paper.qmd
model and paper so far

---
title: "Joe Biden Predicted to Win 59% of the Popular Vote in the 2024 U.S. Presidential Election"
subtitle: "A 2024 U.S. Election Forecast Using Multi-Level Regression with Post-Stratification (MRP)"
author: 
  - Talia Fabregas
  - Fatimah Yunusa
thanks: "Code and data are available at: https://github.com/taliafabs/US-Election-Forecast-2024.git"
date: today
date-format: long
abstract: "The 2024 U.S. Presidential Election will take place on November 5th and it will be a rematch between President Joe Biden and Former President Donald Trump. Using multi-level regression with post-stratification (MRP) we found that ... Based on our results, we can predict that Joe Biden will win 59% of the popular vote. "
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(rstanarm)
library(readr)
library(ggplot2)
library(knitr)
library(arrow)

## load data
survey_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/survey_analysis_data.parquet")

poststrat_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
```

# Introduction

The 2024 U.S. Presidential Election will take place on Tuesday November 5 2024. Incumbent President Joe Biden will seek a second term in a rematch against his 2020 opponent and, arguably, the most polarizing figure in American politics, former President Donald Trump.

Packages used for data cleaning, model, visualizations, and tables: `tidyverse` @citetidyverse, `janitor` @citejanitor, `rstanarm`, `readr`, `ggplot2`, `knitr`, and `arrow`.

The remainder of this paper is structured as follows. @sec-data discusses the survey and post-stratification data used.

Talk about polarization in a discussion point instead of introduction

Clear gap that needs to be filled ... what is the research gap and why is this important?

-   2024 US Presidential Election

-   how consequential this election is

-   what is on the ballot

-   women's rights

-   trump's project 2025 thing

If someone's done it before not on this data set not in this context

Previous groups have looked at ...

@citeR was usd

ESTIMAND: SUPPORT FOR BIDEN

We will use ... to estimate support for President Joe Biden in the upcoming election.

# Data {#sec-data}

```{r}
#| include: false
#| warning: false
#| message: false

library(arrow)
library(dplyr)
survey_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/survey_analysis_data.parquet")
# poststrat_analysis_data <- read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
print(survey_analysis_data)

proportion_biden <- survey_analysis_data %>%
  filter(vote_biden == 1) %>%
  summarise(proportion = n() / nrow(survey_analysis_data))

proportion_biden_by_state <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(proportion_biden = mean(vote_biden == 1, na.rm = TRUE))

proportion_by_state <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(proportion = n() / nrow(survey_analysis_data))

# only includes 7 observations from hawaii
# Hawaii is historically a Democratic stronghold
# only 4 samples from kansas
```

```{r}
#| echo: false
#| warning: false
#| label: fig-survey_presidential_preferences
#| fig-cap: The Distribution of Presidential Preferences, by gender and highest level of education
survey_analysis_data |>
  ggplot(aes(x = education_level, fill = vote24)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    x = "Highest education",
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

## Individual-Level Survey Data {#sec-survey}

Our survey data is from the America's Political Pulse Survey conducted by the Polarization Research Lab. The Polarization Research Lab is a research group founded by top political science research scholars at Dartmouth College, Stanford University, and the University of Pennsylvania, dedicated to applying scientific research methods to the study of democracy and political polarization @americaspoliticalpulse. The Polarization Research Lab studies political polarization across the United States by conducting the Political Pulse Survey each week and making the results available to anyone, in real time, via their website. The America's Political Pulse Survey interviews 1000 American adults each week and asks them questions that aim to track affective polarization, support for violations of democratic norms, and support for the use of political violence in America @americaspoliticalpulse. The Polarization Research Lab makes high-level results of their survey easily accessible and visible via their website's dashboard.

The specific survey data set used is the America's Political Pulse Survey Data Week 3 2024 January 12-19. It contains the responses of the 1000 American adults interviewed between January 12-19 2024. Demographic information about respondents, including sex, age, race, home state, employment status, marital status, and the size and type of city that they live in is included. Participants were also asked who they voted for in the 2016 and 2020 U.S. presidential characteristics, what their party affiliation is, and what their political ideology is.

However, this data set has a few major limitations. It contains very few respondents from Hawaii, Wyoming, and Kansas. It contains zero respondents from Vermont.

The survey does not contain a question about preferred 2024 Presidential Candidate. In fact, it was conducted in January 2024, before Super Tuesday took place and Donald Trump had secured almost enough delegates to win the Republican nomination. At the time when this survey was conducted, former President Donald Trump, former South Carolina Governor Nikki Haley, and Florida Governor Ron DeSantis were all GOP Presidential primary candidates. Media personality Vivek Ramaswamy ended his primary campaign on January 15th, however this was in no way a clear indication that Trump had secured the nomination. As of March 11, 2024, former President Trump has secured 1078 out of the 1215 delegates needed to win the GOP nomination.

Survey Data figures and tables go here

Within our survey data, the vote distribution between Trump and Biden varies by sex and race. As seen in @fig-survey_presidential_preferences1, Black females were more likely to vote for President Joe Biden (D-Delaware) than former President Donald Trump (R-Florida) in 2020, whereas white males were more likely to vote for former President Trump. This indicates that race might be a predictor of who an individual's preferred presidential candidate is.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-racetable
#| tbl-cap: Races of Survey Respondents

races <- survey_analysis_data %>% 
  group_by(races) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "races")

races <- races |>
  select(races, n, pct)

races <- races |>
  rename(Count = n) |>
  rename(Race = races) |>
  rename(Proportion = pct)

kable(races)

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-survey_presidential_preferences1
#| fig-cap: The Distribution of Presidential Preferences, by gender and race

vote2020 <- survey_analysis_data |>
  filter(presvote20post == "Joe Biden" | presvote20post == "Donald Trump")

vote2020 |>
  ggplot(aes(x = races, fill = presvote20post)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    x = "Race",
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

# 73% of respondents are white
# non hispanic white makes up 57.8% of the US population

```

```{r}
# write about America's urban-rural divide
#| echo: false
#| warning: false
#| label: fig-survey_urban_rural
#| fig-cap: The Distribution of Presidential Preferences of Respondents Living in Urban vs Rural Areas
vote2020 |>
  ggplot(aes(x = urban, fill = vote24)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

# do a table of the proportion of respondents for trump vs biden in each state
# overall is like 51% biden
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-popvotesurvey
#| fig-cap: Preferred 2024 Presidential Candidate (Survey Data)

num_respondents <- nrow(survey_analysis_data)
num_biden <- sum(survey_analysis_data$vote_biden)
num_trump <- num_respondents - num_biden

support_table <- tibble(
  num_biden = num_biden,
  prop_biden = num_biden / num_respondents,
  num_trump = num_trump,
  prop_trump = num_trump / num_respondents
)

kable(support_table)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-electoral-map-survey
#| fig-cap: Electoral map based solely on the survey data
library(ggplot2)
library(statebins)
survey_electoral_map <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(prop = mean(vote_biden))

survey_electoral_map %>%
  mutate(statename = str_to_title(state)) %>%
  ggplot(aes(fill = prop, state = statename)) +
  geom_statebins() +
  scale_fill_gradient2(low = "#d12531", high = "#244999",
                       mid = "white", midpoint = mean(survey_electoral_map$prop)) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(fill = "Proportion Voting \nfor Biden",
       title = "Map of the US States that plan to support Biden or Trump")

# note extremely small sample sizes from Hawaii, Kansas, and Wyoming
# note that survey sample size was only 1000
```

@fig-electoral-map-survey illustrates the proportion of voters in each U.S. state who plan to support President Biden in the 2024 Presidential Election. We can see that there is strong support for President Biden in Washington, Oregon, New York, New Jersey, Maryland, Massachusetts, and Connecticut. We also see that Oklahoma, Idaho, and Utah heavily favor former President Trump. These results are expected and reflective of the 2020 election. These states have also been Dem and GOP strongholds since 1992 (or whatever year).

```{r}
#| echo: false
#| warning: false
#| message: false
#| table: tbl-electoralcollegesurvey
#| fig-cap: Electoral College based on survey data

electoral_college <- read_csv("electoral_college_votes.csv", show_col_types = FALSE)
# downloaded and cleaned it
electoral_college$state <- tolower(electoral_college$state)

electoral_college$electoral_votes <- as.numeric(electoral_college$electoral_votes)

library(dplyr)

prop_biden_by_state <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(prop = mean(vote_biden))

prop_biden_by_state$state <- tolower(prop_biden_by_state$state)

merged_survey_ec <- merge(electoral_college, prop_biden_by_state, by="state")

biden_states<- merged_survey_ec |>
  filter(prop >= 0.5)

trump_states <- merged_survey_ec |>
  filter(prop < 0.5)

biden_ec_survey <- sum(biden_states$electoral_votes)
trump_ec_survey <- sum(trump_states$electoral_votes)

electoral_college_survey <- tibble(
  Biden = biden_ec_survey,
  Trump = trump_ec_survey
)

kable(electoral_college_survey)
```

However, this electoral map also highlights some of the limitations fo the survey data. Wyoming and Kansas are very blue in the electoral map based solely on the survey data, however that is very far from recent election results in those states. In fact, Wyoming was the most Pro-Trump state in both the 2016 and 2020 presidential elections, with Trump receiving 2016% and 2020% of the vote in each of those elections, respectively. The survey data set includes very few examples from small states.

There are no respondents from Vermont in the survey data, thus Vermont is missing from the electoral college map and @tbl-electoralcollegesurvey is missing the 3 electoral votes allocated to Vermont.

## Post Stratification Data {#sec-poststrat}

The post-stratification data used is comes from the American Community Survey (ACS) 2022 provided by IPUMS ( @citeIPUMS ).

Talk more about it.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-racetable2
#| tbl-cap: Races of Population Data

races2 <- poststrat_analysis_data %>% 
  group_by(races) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "races")

races2 <- races2 |>
  select(races, n, pct)

races2 <- races2 |>
  rename(Count = n) |>
  rename(Race = races) |>
  rename(Proportion = pct)

kable(races2)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-demographics1
#| fig-cap: Voter Demographics in Survey and Post-Stratification Data Sets
gender <- survey_analysis_data %>% 
  group_by(sex) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "gender") %>% 
  rename(level = sex)

races <- survey_analysis_data %>% 
  group_by(races) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "races") %>% 
  rename(level = races)

education <- survey_analysis_data %>% 
  group_by(education_level) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "education") %>% 
  drop_na(education_level) %>% 
  rename(level = education_level)

age <- survey_analysis_data %>% 
  group_by(age_bracket) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "age") %>% 
  rename(level = age_bracket)

hispanic <- survey_analysis_data %>% 
  group_by(race_hispanic) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "hispanic") %>% 
   rename(level = race_hispanic)

gender_poststrat <- poststrat_analysis_data %>% 
  group_by(sex) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "gender") %>% 
  rename(level = sex)

races_poststrat <- poststrat_analysis_data %>% 
  group_by(races) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "races") %>% 
  rename(level = races)

education_poststrat <- poststrat_analysis_data %>% 
  group_by(education_level) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "education") %>% 
  drop_na(education_level) %>% 
  rename(level = education_level)

age_poststrat <- poststrat_analysis_data %>% 
  group_by(age_bracket) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "age") %>% 
  rename(level = age_bracket)

state_poststrat <- poststrat_analysis_data %>% 
  group_by(state) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "state") %>% 
   rename(level = state)

hispanic_poststrat <- poststrat_analysis_data %>% 
  group_by(race_hispanic) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "hispanic") %>% 
   rename(level = race_hispanic)

## combine the demographic variables into a single data set

demographic_variables <- rbind(gender, gender_poststrat, 
                               age, age_poststrat, 
                               races, races_poststrat,
                               hispanic, hispanic_poststrat,
                               education, education_poststrat)

demographic_variables |> ggplot(aes(as.factor(level), pct, group=as.factor(type), linetype = as.factor(type))) +
  geom_line() + facet_grid(~group, scales = "free") +
  theme(axis.text.x = element_text(angle=70, size = 6, hjust = 1), legend.position = "bottom") +
  labs(x = "Category", y = "Proportion", linetype = "data set",
       title = "Voter Demographics") +
  scale_y_continuous(labels = scales::percent) +
  geom_hline(yintercept = 0.5, alpha = 0.35)

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-stateprops
#| tbl-cap: Survey and Post-Stratification Data Proportion of Voters by State

survey_analysis_data$state <- tolower(survey_analysis_data$state)
state <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(n = n()) %>%
  mutate(pct = n/sum(n), type = "survey",
         group = "state") %>%
   rename(level = state)

state_poststrat <- poststrat_analysis_data %>%
  group_by(state) %>%
  summarise(n = n()) %>%
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "state") %>%
   rename(level = state)

voters_by_state <- rbind(state, state_poststrat)

voters_by_state |> ggplot(aes(as.factor(level), pct, group=as.factor(type), linetype = as.factor(type))) +
  geom_line() + facet_grid(~group, scales = "free") +
  theme(axis.text.x = element_text(angle=70, size = 6, hjust = 1)) +
  labs(x = "State", y = "Proportion", linetype = "Data Set")+
       # title = "Proportion of Voters by State") +
  scale_y_continuous(labels = scales::percent)
```

Talk way more about it.

\newpage

# Model

We will use multilevel regression with post-stratification (MRP) for our analysis. MRP is ...

Steps for MRP are:

## Model set-up

To predict whether a person will voter for Joe Biden or Donald Trump, we built a logistic regression model using data from the America's Political Pulse Survey [@americaspoliticalpulse] and post-stratified it using ACS census data [@citeIPUMS]. Logistic regression only works for binary classification, therefore we created a vote_biden variable during our data cleaning process which is equal to 1 if the person will vote for Joe Biden and 0 if the person will vote for Donald Trump.

The logistic regression model used to predict support for Biden using sex, age, race, highest level of education, state, and whether the individual lives in an urban or rural area is as follows:

<small>

$$
y_i = \beta_0 + \beta_1 \times \text{sex}_i + \beta_2 \times \text{age\_bracket}_i + \beta_3 \times \text{race}_i + \beta_4 \times \text{education\_level}_i + \beta_5 \times \text{state}_i + \beta_6 \times \text{urban}_i 
$$

</small>

logistic activation <smalL> $$
Pr(vote\_biden=1) = logit^{-1}(\beta_0 + \beta_1 \times \text{sex}_i + \beta_2 \times \text{age\_bracket}_i + \beta_3 \times \text{race}_i + \beta_4 \times \text{education\_level}_i + \beta_5 \times \text{state}_i + \beta_6 \times \text{urban}_i )
$$ </small>

another try <small>

$$
Pr(vote\_biden=1) = \text{logit}^{-1}(\beta_0 + \beta_1 \text{sex}_i + \beta_2  \text{age\_bracket}_i + \beta_3 \text{race}_i + \beta_4 \text{education\_level}_i + \beta_5 \text{state}_i + \beta_6  \text{urban}_i )
$$

</small>

We run our model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

We use the `predict()` function in R @citeR to apply our logistic regression model to the ACS Census data @citeIPUMS.

Model weaknesses: logistic regression output is binary so it does not include the options of not voting or voting for a third-party candidate.

Concerns about model fit

### Model justification

We expect to see increased support for President Biden amongst Black college-educated women in urban areas.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# overall support
# support by state
# electoral college map

library(rstanarm)

# load the model with readRDS
presidential_vote_model <-
  readRDS(file = here::here("models/presidential_vote_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelsummary
#| tbl-cap: "Coefficient estimates"
#| warning: false

# looking at coefficient estimates using modelsummary()
modelsummary::modelsummary(presidential_vote_model)

```

# Results

```{r}
#| eval: true
#| echo: false
#| warning: false
#|abel: tbl-supporttbl
#| tbl-cap: "Estimates for support for Biden in each state"

library(arrow)
library(broom.mixed)
library(gutenbergr)
library(haven)
library(knitr)
library(labelled)
library(modelsummary)
library(rstanarm)
library(tidybayes)
library(tidyverse)

# get proportions by demographic, then find proportion of each cell relative to population
props <- poststrat_analysis_data %>%
  group_by(state, races, age_bracket, sex, education_level, urban) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  group_by(state) %>%
  mutate(prop = n / sum(n))

poststrat_data_cells <-
  poststrat_analysis_data |>
  count(state, sex, age_bracket, races, education_level, urban)

poststrat_data_cells <-
  poststrat_data_cells |>
  mutate(prop = n / sum(n),
         .by = state)

poststrat_data_cells <- poststrat_data_cells %>%
  mutate(state = tolower(state))

biden_support_by_state <-
  presidential_vote_model |>
  add_epred_draws(newdata = poststrat_data_cells) |>
  rename(support_biden_predict = .epred) |>
  mutate(support_biden_predict_prop = support_biden_predict * prop) |>
  ungroup() |>
  summarise(support_biden_predict = sum(support_biden_predict_prop),
            .by = c(state, .draw)) |>
  summarise(
    mean = mean(support_biden_predict),
    lower = quantile(support_biden_predict, 0.025),
    upper = quantile(support_biden_predict, 0.975),
    .by = state
  )

kable(biden_support_by_state)
```

electoral college

```{r}
# googled the electoral college and made a csv
electoral_college <- read_csv("electoral_college_votes.csv", show_col_types = FALSE)
# downloaded and cleaned it

electoral_college$electoral_votes <- as.numeric(electoral_college$electoral_votes)

merged_ec <- merge(biden_support_by_state, electoral_college, by = "state")

biden_states_mean <- merged_ec |>
  filter(mean >= 0.5)

trump_states_mean <- merged_ec |>
  filter(mean < 0.5)

biden_ec_mean <- sum(biden_states_mean$electoral_votes)
trump_ec_mean <- sum(trump_states_mean$electoral_votes)

biden_states_mean <- merged_ec |>
  filter(lower >= 0.5)

trump_states_mean <- merged_ec |>
  filter(lower < 0.5)

biden_ec_lower<- sum(biden_states_mean$electoral_votes)
trump_ec_lower <- sum(trump_states_mean$electoral_votes)

```

The estimates for the support for Biden in each state are shown in @tbl-supporttbl

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| #| label: tbl-popvotepost
#| tbl-cap: "Estimated Popular Vote"
biden_support_overall <- presidential_vote_model %>%
  add_epred_draws(newdata = poststrat_data_cells) %>%
  rename(support_biden_predict = .epred) %>%
  mutate(support_biden_predict_prop = support_biden_predict * prop) %>%
  ungroup() %>%
  summarise(
    "Lower Estimate" = quantile(support_biden_predict, 0.05),
    "Mean Estimate" = mean(support_biden_predict),
    "Upper Estimate" = quantile(support_biden_predict, 0.95)
  )

kable(biden_support_overall)
```

```{r}
#| eval: true
#| echo: false
#| warning: false
#|abel: tbl-supporttbl
#| tbl-cap: "Estimates for support for Biden in each state"

library(arrow)
library(broom.mixed)
library(gutenbergr)
library(haven)
library(knitr)
library(labelled)
library(modelsummary)
library(rstanarm)
library(tidybayes)
library(tidyverse)

# get proportions by demographic, then find proportion of each cell relative to population
props <- poststrat_analysis_data %>%
  group_by(state, races, age_bracket, sex, education_level, urban) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  group_by(state) %>%
  mutate(prop = n / sum(n))

poststrat_data_cells <-
  poststrat_analysis_data |>
  count(state, sex, age_bracket, races, education_level, urban)

poststrat_data_cells <-
  poststrat_data_cells |>
  mutate(prop = n / sum(n),
         .by = state)

poststrat_data_cells <- poststrat_data_cells %>%
  mutate(state = tolower(state))

biden_support_by_state <-
  presidential_vote_model |>
  add_epred_draws(newdata = poststrat_data_cells) |>
  rename(support_biden_predict = .epred) |>
  mutate(support_biden_predict_prop = support_biden_predict * prop) |>
  ungroup() |>
  summarise(support_biden_predict = sum(support_biden_predict_prop),
            .by = c(state, .draw)) |>
  summarise(
    mean = mean(support_biden_predict),
    lower = quantile(support_biden_predict, 0.025),
    upper = quantile(support_biden_predict, 0.975),
    .by = state
  )

kable(biden_support_by_state)
```

visualization here

```{r}
#| echo: false
#| warning: false
#| eval: true
#| label: fig-estimatedpropsupport
#| fig-cap: "Estimated proportion of each state voting for Biden in 2024 (Post-Stratification)"

library(scales)  
biden_support_by_state |>
  ggplot(aes(y = mean, x = fct_reorder(state, mean))) +
  geom_point(color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0, color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  labs(
    x = "State",
    y = "Estimated proportion support for Biden"
  ) +
  theme_classic() +
  scale_y_continuous(labels = scales::percent)+  # Setting y-axis to display percentages
  coord_flip() +
  theme(axis.text.y = element_text(size = 6)) 
```

compared to survey data shown in @fig-estimatedpropsupport2

```{r}
#| echo: false
#| eval: true
#| warning: false
#| label: fig-estimatedpropsupport2
#| fig-cap: "Estimated proportion of each state voting for Biden in 2024 Post-Stratification vs Survey Data"

survey_analysis_data$state <- tolower(survey_analysis_data$state)
biden_support_by_state |>
  ggplot(aes(y = mean, x = fct_reorder(state, mean))) +
  geom_point(color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0, color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
  geom_point(
    data = survey_analysis_data |>
      summarise(n = n(),
                .by = c(state, vote_biden)) |>
      mutate(prop = n / sum(n),
             .by = state) |>
      filter(vote_biden == 1),
    aes(y = prop, x = state),
    color="#A9A9A9"
    ) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  labs(
    x = "State",
    y = "Estimated proportion support for Biden"
  ) +
  theme_classic() +
  scale_y_continuous(labels = scales::percent)+ # Setting y-axis to display percentages
  coord_flip() +
  theme(axis.text.y = element_text(size = 5)) 
```

electoral map

```{r}
#| echo: false
#| eval: true
#| warning: false
#| label: fig-electoralmappoststrat
#| fig-cap: "Electoral map based on post-stratification data"
library(ggplot2)
library(statebins)
survey_electoral_map <- biden_support_by_state %>%
  mutate(prop = mean)

survey_electoral_map %>%
  mutate(statename = str_to_title(state)) %>%
  ggplot(aes(fill = prop, state = statename)) +
  geom_statebins() +
  scale_fill_gradient2(low = "#d12531", high = "#244999",
                       mid = "white", midpoint = mean(survey_electoral_map$prop)) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(fill = "Proportion Voting \nfor Biden",
       title = "Map of the US States that plan to support Biden or Trump (Post-Stratification Data)")
```

type here

# Discussion

## Popular Vote Projection {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Electoral College Projection

## Polarization and America's Urban-Rural Divide

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

## Data Cleaning

# Model details {#sec-model-details}

type something here

```{r}
#| eval: true
#| echo: false
#| warning: false
#| label: fig-coefficientestimatesplot
#| layout-ncol: 2
#| fig-cap: "Coefficient Estimate"

# coefficients|> ggplot(aes(estimate, term)) +
#   geom_point() +
#   geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) +
#   labs(title = "Coefficient Estimates",
#        x = "Estimate", y = "Coefficient") +
#   theme(axis.text.y = element_text(size = 5))
```

## Correlation Map

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

# pp_check(first_model) +
#   theme_classic() +
#   theme(legend.position = "bottom")
# 
# posterior_vs_prior(first_model) +
#   theme_minimal() +
#   scale_color_brewer(palette = "Set1") +
#   theme(legend.position = "bottom") +
#   coord_flip()
```

Model plot (still need to fix this)

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelplot1
#| tbl-cap: "Model Plot"
#| warning: false

# modelplot(presidential_vote_model, conf_level = 0.9) +
#   labs(x = "90 per cent credibility interval")

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2
# 
# plot(first_model, "trace")
# 
# plot(first_model, "rhat")
```

\newpage

# References