### 2024-03-17 paper/paper.qmd
final edits

---
title: "Joe Biden Predicted to Win 56.59% of the Popular Vote in the 2024 U.S. Presidential Election"
author: 
  - Talia Fabregas
  - Fatimah Yunusa
  - Aamishi Sundeep Avarsekar
thanks: "Code and data are available at: https://github.com/taliafabs/US-Election-Forecast-2024.git"
date: today
date-format: long
abstract: "The 2024 U.S. Presidential Election will take place on November 5th and it will be a rematch between President Joe Biden and Former President Donald Trump. In this report, we used a logistic regression model along with multi-level post-stratification (MRP) to predict the winner of the upcoming election. Using the results of our model, we predict that President Biden will win 56.59% of the popular vote and defeat former President Trump 467 to 71 in the electoral college, and win a second term in the White House. However, due to the limited nature of our survey data, the error range of our electoral college prediction is large."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#| eval: true

library(tidyverse)
library(janitor)
library(rstanarm)
library(readr)
library(ggplot2)
library(knitr)
library(arrow)
library(kableExtra)
library(modelsummary)
library(MASS)
library(broom.mixed)
library(gutenbergr)
library(haven)
library(labelled)
library(tidybayes)
library(tidyverse)

## load data
# change file path to the location of cleaned survey and poststrat data on local computer
survey_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/survey_analysis_data.parquet")

poststrat_analysis_data <- arrow::read_parquet("/Users/talia/US-Election-Forecast/data/analysis_data/poststrat_analysis_data.parquet")
```

# Introduction

The U.S. Presidential election will take place on Tuesday, November 5th, 2024, as America faces unprecedented levels of political polarization and diminished levels of trust in democratic institutions. As of March 11, 2024, former President Trump has 1078 out of the 1215 delegates needed to win the Republican nomination and President Biden has 1872 out of the 1968 delegates needed to win the Democratic nomination, therefore this election will feature the same candidates as the 2020 election. President Joe Biden will seek a second term and former President Donald Trump will try to become the second president to serve two non-consecutive terms. The only President to serve two non-consecutive terms was Grover Cleveland from 1885 to 1889 and 1893 to 1897 [@citeWaxman]. The 2024 election will define the future of women’s rights, LGBTQ+ rights, the separation of church and state, immigration policy, and foreign policy in America. America is at a turning point: Trump and Biden have very different ideologies, policy ideas, and visions for the future of the nation. Former President Trump has celebrated that he was “able to kill Roe v. Wade” [@citeKapur]. With 2022 Supreme Court decision to overturn Roe v. Wade, there will be nothing to stop Trump from signing a bill to restrict access to, or even criminalize, abortion, contraception, or IVF across the United States. In February 2024, Trump stated that he “would encourage” the Putin regime to do “whatever the hell they want” to NATO allies who don’t meet their defense spending targets [@citeSullivan]. On the contrary, President Biden has stood for women’s rights, protecting same-sex marriage, pro-immigration policies, and with Ukraine in their fight for freedom following Russia’s invasion. The January 6th, 2021, insurrection at the U.S. Capitol was a defining moment in the current era of political polarization in America. Trump refused to acknowledge his loss to Joe Biden in the 2020 U.S. presidential election. In August 2020, Trump claimed that the only way he was going to lose the election was “if the election is rigged” [@citeAxelrod]. Since November 2020, he has falsely claimed that the election was fraudulent, and that Biden stole the election from him. On January 6, 2021, an angry pro-Trump mob stormed the U.S. Capitol in an attempt to "stop the steal," or in other words, to stop the joint session of congress from certifying Biden’s electoral college victory [@citeLonsdorf]. It is unclear whether former President Trump will concede if he is defeated again by President Biden in the 2024 election.

Our survey data set was provided by the Polarization Research Lab. The Polarization Research Lab was formed by a group of researchers to increase understanding of where partisan animosity comes from and what can be done to help [@americaspoliticalpulse]. Our post-stratification data set was provided by the Integrated Public Use Microdata Series (IPUMS). IPUMS is an organization which lays out survey and census data with the help of 105 statistical agencies [@citeIPUMS]. We used the data from the 2022 American Community Survey (ACS) for this report [@citeIPUMS]. We use multi-level regression with post-stratification (MRP) to predict the results of the 2024 U.S. Presidential Election. MRP uses a smaller survey data set to fit a model to predict vote preference. We built a logistic regression model, using sex, age, race, hispanic, highest level of education, and whether the respondent lives in an urban or rural area as predictors to estimate support for President Biden. We fit this model using our survey data set, and apply it to our post-stratification data set to predict the popular vote and electoral college results of the 2024 election.

Our report has 4 sections. In @sec-data, we examine the data sets that we used for our report and present how variables in our survey data set interact as well as summary statistics. In @sec-model, we outline how we built a logistic regression model and used MRP to predict the results of the 2024 U.S. presidential election. In @sec-results, we present the findings of our model, which include a prediction for the popular vote and electoral college results. In @sec-discussion, we discuss what we can learn from our results, how the weaknesses of our survey data set translate into model weaknesses and limitations, and how this report can be extended and improved.

We used R [@citeR], and the `tidyverse` [@citetidyverse], `janitor` [@citejanitor], `ggplot` [@citeggplot] packages to clean our survey and post-stratification data sets, fit our model, apply our model, and create our data visualizations.

# Data {#sec-data}

Our survey data is from the America's Political Pulse Survey conducted by the Polarization Research Lab. The Polarization Research Lab is a research group founded by top political science research scholars at Dartmouth College, Stanford University, and the University of Pennsylvania, dedicated to applying scientific research methods to the study of democracy and political polarization @americaspoliticalpulse. The America's Political Pulse Survey interviews 1000 American adults each week and asks them questions that aim to track affective polarization, support for violations of democratic norms, and support for the use of political violence in America @americaspoliticalpulse. High-level survey results are available on the Polarization Research Lab website.

Our post-stratification data is a subset of the 2022 American Communities Survey (ACS) provided by the Integrated Public Use Microdata Series [@citeIPUMS].

## Survey Data {#sec-survey}

We selected the America's Political Pulse Survey Data Week 3 2024 from the Polarization Research Lab as our survey data. It contains the responses of the 1000 American adults interviewed between January 12-19 2024. Samples were collected via interviews. The population of the survey, the respondents, are paid survey-takers from the YouGov survey platform @americaspoliticalpulse. Demographic information about respondents, including sex, age, race, home state, employment status, marital status, and the size and type of city that they live in is included. Participants were also asked questions about who they voted for in the 2016 and 2020 U.S. presidential elections, stances on political violence, what their party affiliation is, the strength of their party affiliation, and what their political ideology is. While the America's Political Pulse survey aims to learn about affective polarization and respect for democratic norms by asking respondents questions regarding their feelings about the importance of voting, towards the Democratic Party and the Republican Party, faith in democratic institutions, and about the use of political violence, we are focusing on demographics and political preferences. As expected, responses vary by ideology and political affiliation.

We are focusing on demographics and other variables that might indicate who each respondent will vote for in the 2024 presidential election. The variables from the America's Political Pulse Survey that we selected include: `pid7`: the respondent's party affiliation and the strength of it; `presvote16post`: who the respondent voted for in the 2016 U.S. presidential election; `pres20votepost`: who the respondent voted for in the 2020 U.S. presidential election; `gender`: the gender of the respondent; `age`: the respondent's age as of January 2024; `race`: respondent's race or ethnicity; `educ`: highest level of education completed; `inputstate`: state of residence; `urbancity2`: whether or not the respondent lives in a city, and if so, the size of that city.

The survey does not contain a question about a preferred 2024 Presidential Candidate. In fact, it was conducted in January 2024, before Super Tuesday took place and Donald Trump had secured almost enough delegates to win the Republican (GOP) nomination. We created a variable `vote_biden`, which is equal to 1 if a respondent's preferred 2024 presidential candidate is Joe Biden, and 0 if it is Donald Trump. The data cleaning process used for the creation of the `vote_biden` variable is outlined in @sec-datacleaning.

In @fig-survey_presidential_preferences1, we can see that support for Biden within our survey data varies by race and sex. Black women showed overwhelming support for President Biden's upcoming re-election campaign, while white men appear to be split between Biden and Trump. Among white survey respondents (69% of all respondents), women were more likely to support President Biden and men were more likely to support Trump. We used code from Chapter 13 of Telling Stories with Data and modified it to produce the bar graphs in this section [@tswd].

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-survey_presidential_preferences1
#| fig-cap: Preferred presidential candidates of survey respondents, by gender and race

survey_analysis_data |>
  ggplot(aes(x = races, fill = vote24)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    x = "Race",
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

We see differences in vote preference among survey respondents with different levels of education. Male voters with up to a high school education or some college education were more likely to prefer Trump than their 4-year college or post-graduate educated counterparts. Only male voters whose highest level of education is a 4-year college degree favored Biden over Trump, while males with a post-graduate education were split. Females with a post-graduate education heavily favored Biden, further highlighting the gender-gap in support for Biden and Trump. This makes sense, as Trump is strongly opposed to abortion rights and since the Supreme Court of the United States overturned Roe v. Wade in June 2022, there is nothing to stop him from signing an outright ban on abortion, contraceptives, or fertility treatments if elected. Females with a 4-year college or post-grad education were more likely to prefer Biden than those with only a high school education.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-survey_presidential_preferences
#| fig-cap: The Distribution of Presidential Preferences, by gender and highest level of education

survey_analysis_data |>
  ggplot(aes(x = education_level, fill = vote24)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    x = "Highest education",
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
# code for this graph is from telling stories with data chapter 13

```

Since the 1990s, America has developed an urban-rural ideological divide and it has kept growing [@citecornellians]. We created a new binary variable, `urban`, indicates whether the respondent lives in an urban or rural area. The data cleaning steps that we used to create this variable are outlined in @sec-datacleaning. First, we added the `urban` variable to our survey data. It is equal to "urban" if `urbancity2` indicates that the respondent lives in a big city, a smaller city, or a suburb, and "rural" if the respondent lives in a small town or rural area. This effectively distinguishes city and suburban voters from small town and rural voters, making our analysis simpler. We found that female survey respondents living in urban areas heavily favored Biden, while male survey respondents living in rural areas heavily favored Trump. In @fig-survey_urban_rural, we can see that while female survey respondents living in rural areas were split and male survey respondents living in urban areas favored Biden by a smaller margin, females in urban areas heavily favored Biden and males in rural areas heavily favored Trump.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-survey_urban_rural
#| fig-cap: Preferred presidential candidate of respondents living in urban vs rural areas

survey_analysis_data |>
  ggplot(aes(x = urban, fill = vote24)) +
  stat_count(position = "dodge") +
  facet_wrap(facets = vars(sex)) +
  theme_minimal() +
  labs(
    y = "Number of respondents",
    fill = "Preferred Presidential Candidate"
  ) +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

# code for this graph is from telling stories with data chapter 13

```

Overall, Biden outperformed Trump in urban areas, with the support of 60.22% of survey respondents living in urban areas. This is unsurprising, and it is a continuation of the pattern observed in recent presidential elections. @citeScala found that both former President Barack Obama and 2016 Democratic nominee Hillary Clinton performed well in densely populated urban areas, but faced increasingly challenging political climates and diminished voter support in rural areas. Hillary Clinton saw a particularly pronounced decline in support in rural areas in the 2016 election, where she was defeated by Donald Trump (Scala, 2016). Our survey data is somewhat consistent with this, as Biden trails Trump among survey respondents living in rural areas, with 46.20% support.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: tbl-survey_urban_rural
#| tbl-cap: Presidential preferences of respondents living in urban and rural areas

urban <- survey_analysis_data |> filter(urban == "urban")
rural <- survey_analysis_data |> filter(urban == "rural")

biden_urban <- round(mean(urban$vote_biden)*100,2)
biden_rural <- round(mean(rural$vote_biden)*100,2)

urban_rural <- tibble(
  "Urban or Rural:" = c("Urban", "Rural"),
  "Biden %" = c(biden_urban, biden_rural),
  "Trump %" = c(100-biden_urban, 100-biden_rural)
)

# Set row names
rownames(urban_rural) <- NULL

# Render the table using kable
kable(urban_rural, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

Using only the survey data, we can calculate the overall support for Biden and the support for Biden in each state to make a popular vote and electoral college prediction. Based on our earlier findings, we believe that being non-white, female, college or post-grad educated, and living in an urban area makes an individual more likely to support Biden. To calculate a popular vote prediction for President Biden based on the survey data, we simply calculated the mean of the `vote_biden` binary indicator variable, and to calculate an electoral college prediction, we added up the electoral college votes of the states where Biden received more than 50% support from survey respondents. We used the steps and code provided by @citeMatthew and the `statebins` package [@statebins] to produce @fig-electoral-map-survey.

```{r}
#| echo: false
#| warning: false
#| eval: true
#| message: false
#| label: fig-electoral-map-survey
#| fig-cap: Electoral college map based on the survey data

library(ggplot2)
library(statebins)

# this code comes from matthewwankiewicz 
# US_election_forecast/paper/forecasting_us_election.rmd
# also cited in text

survey_electoral_map <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(prop = mean(vote_biden))

survey_electoral_map %>%
  mutate(statename = str_to_title(state)) %>%
  ggplot(aes(fill = prop, state = statename)) +
  geom_statebins() +
  scale_fill_gradient2(low = "#d12531", high = "#244999",
                       mid = "white", midpoint = mean(survey_electoral_map$prop)) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(fill = "Proportion Voting \nfor Biden")

# note extremely small sample sizes from Hawaii, Kansas, and Wyoming
# note that survey sample size was only 1000
```

@fig-electoral-map-survey illustrates the proportion of voters in each state who plan to support President Biden in the 2024 election. We can see that there is very strong support for President Biden in "deep blue" states such as Massachusetts, Connecticut, New Jersey, and Oregon. A majority of respondents in New York, New Jersey, Maryland, and California also support President Biden. On the contrary, Oklahoma, Idaho, Utah, and Alaska appear to be "deep red" states, which indicates that a high proportion of survey respondents from those states support former President Trump. Notably, Wyoming appears to be dark blue, indicating that 100% of survey respondents from that state support President Biden. However, Wyoming is a Republican stronghold, and the only Democratic presidential candidate who has won it since 1948 is Lyndon B. Johnson in 1964 [@citeCNN]. Wyoming was the most pro-Trump state in both the 2016 and 2020 presidential elections. Trump outperformed 2016 Democratic nominee Hillary Clinton and President Joe Biden in Wyoming by 46 and 43.3 percentage points, respectively [@citeCNN]. Our survey data contains one respondent from Wyoming, who leans Democrat and voted for Hillary Clinton in 2016 and Joe Biden in 2020. There are over half a million people in Wyoming, and @fig-electoral-map-survey reflects the political preferences of exactly one of them. Similarly, Hawaii (7 respondents) appears to be a "red" state, indicating that the majority of survey respondents from that state prefer former President Trump. However, Hawaii is a Democratic stronghold and it has only been carried by a Republican presidential candidate twice since it became a state in 1959 [@citeCNN]. President Biden and Hillary Clinton won 63.7% and 62.2% of the vote in Hawaii in 2020 and 2016, respectively [@citeCNN].

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-surveyresults
#| tbl-cap: Popular vote and electoral college based on survey data

electoral_college <- read_csv("electoral_college_votes.csv", show_col_types = FALSE)
# downloaded and cleaned it
electoral_college$state <- tolower(electoral_college$state)

electoral_college$electoral_votes <- as.numeric(electoral_college$electoral_votes)

library(dplyr)

prop_biden_by_state <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(prop = mean(vote_biden))

prop_biden_by_state$state <- tolower(prop_biden_by_state$state)

merged_survey_ec <- merge(electoral_college, prop_biden_by_state, by="state")

biden_states<- merged_survey_ec |>
  filter(prop >= 0.5)

trump_states <- merged_survey_ec |>
  filter(prop < 0.5)

biden_ec_survey <- sum(biden_states$electoral_votes)
trump_ec_survey <- sum(trump_states$electoral_votes)

num_respondents <- nrow(survey_analysis_data)
num_biden <- sum(survey_analysis_data$vote_biden)
num_trump <- num_respondents - num_biden

percent_biden = round((num_biden / num_respondents)*100, 2)
percent_trump = round((num_trump / num_respondents)*100,2)

biden_survey <- tibble(
  "Survey Estimate:" = c("Num Votes", "% Votes", "Electoral College"),
  "Biden" = c(num_biden,
              percent_biden,
              biden_ec_survey
              ),
  "Trump" = c(num_trump, 
              percent_trump, 
              trump_ec_survey)
)

# Set row names
rownames(biden_survey) <- NULL

# Render the table using kable
kable(biden_survey, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

@tbl-surveyresults summarizes our predictions for the popular vote and electoral college based on our survey data. We can use this to predict that Joe Biden will win 51.84% of the popular vote, and win re-election by defeating Donald Trump 324 to 211 in the electoral college.

Our surey data set contains zero respondents from Vermont, so it is not shown @fig-electoral-map-survey and the number of electoral colleges in @tbl-surveyresults adds up to 535 instead of 538 because Vermont's 3 electoral votes were not included. The absence of any respondent from Vermont and the fact that there are very few responses from Kansas, Hawaii, Rhode Island, and Wyoming may affect the accuracy of this prediction. This is a significant weakness, and we will discuss it in more depth, as well as how it contributes to the weaknesses and limitations of our model in @sec-weaknesses-limitations.

## Post Stratification Data {#sec-poststrat}

Our post-stratification data is a subset of the American Community Survey (ACS) 2022 provided by IPUMS [@citeIPUMS]. We visited the IPUMS data base, selected the ACS 2022, selected relevant variables, and downloaded the data set. Our raw post-stratification data set has 500,000 records. The variables that we selected mirror the ones that we have in our survey data set: `age`: the age of the census respondent; `sex`: only male and female are considered, for simplicity and to align with our survey data; `race`: the race of the census respondent; `hispan`: whether the respondent is of Hispanic origin or not; `educ`: highest level of education completed; `stateicp`: state of residence; `metro`: whether the respondent lives in a metropolitan area or not.

The sex, race, educ, and metro variables were cleaned and re-factored to mirror our survey data and to allow us to apply our model to our post-stratification data. @fig-demographics1 shows the demographic statistics of our survey and post-stratification data sets. The steps and code for producing @fig-demographics1 were obtained from @citeMatthew. We also created a `race_hispanic`, which is equal to either "hispanic" or "not hispanic" to ensure that our survey and post-stratification data sets would align, because this was a separate variable from race in our post-stratification data. The data cleaning steps used to create the `urban` variable to make our post-stratification data set mirror our survey data set are outlined in @sec-datacleaning.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-demographics1
#| fig-cap: Survey vs post-stratification voter demographics

# steps here were obtained from matthewwankiewicz US_election_forecast github repo

gender <- survey_analysis_data %>% 
  group_by(sex) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "gender") %>% 
  rename(level = sex)

races <- survey_analysis_data %>% 
  group_by(races) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "races") %>% 
  rename(level = races)

education <- survey_analysis_data %>% 
  group_by(education_level) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "education") %>% 
  drop_na(education_level) %>% 
  rename(level = education_level)

age <- survey_analysis_data %>% 
  group_by(age_bracket) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "survey",
         group = "age") %>% 
  rename(level = age_bracket)

hispanic <- survey_analysis_data %>%
  group_by(race_hispanic) %>%
  summarise(n = n()) %>%
  mutate(pct = n/sum(n), type = "survey",
         group = "hispanic") %>%
   rename(level = race_hispanic)

gender_poststrat <- poststrat_analysis_data %>% 
  group_by(sex) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "gender") %>% 
  rename(level = sex)

races_poststrat <- poststrat_analysis_data %>% 
  group_by(races) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "races") %>% 
  rename(level = races)

education_poststrat <- poststrat_analysis_data %>% 
  group_by(education_level) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "education") %>% 
  drop_na(education_level) %>% 
  rename(level = education_level)

age_poststrat <- poststrat_analysis_data %>% 
  group_by(age_bracket) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "age") %>% 
  rename(level = age_bracket)

state_poststrat <- poststrat_analysis_data %>% 
  group_by(state) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "state") %>% 
   rename(level = state)

hispanic_poststrat <- poststrat_analysis_data %>%
  group_by(race_hispanic) %>%
  summarise(n = n()) %>%
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "hispanic") %>%
rename(level = race_hispanic)

## combine the demographic variables into a single data set

demographic_variables <- rbind(gender, gender_poststrat, 
                               age, age_poststrat, 
                               education, education_poststrat)

demographic_variables |> ggplot(aes(as.factor(level), pct, group=as.factor(type), linetype = as.factor(type))) +
  geom_line() + facet_grid(~group, scales = "free") +
  theme(axis.text.x = element_text(angle=70, size = 6, hjust = 1), legend.position = "bottom") +
  labs(x = "Category", y = "Proportion", linetype = "data set") +
  scale_y_continuous(labels = scales::percent) +
  geom_hline(yintercept = 0.5, alpha = 0.35) 
```

@fig-datasetraces illustrates the demographic patterns in our post-stratification and survey data sets. The steps and code for producing @fig-stateprops were obtained from @citeMatthew.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-datasetraces
#| fig-cap: Survey vs post-stratification voter race demographics
race_variables <- rbind(races, races_poststrat,
                        hispanic, hispanic_poststrat)

race_variables |> ggplot(aes(as.factor(level), pct, group=as.factor(type), linetype = as.factor(type))) +
  geom_line() + facet_grid(~group, scales = "free") +
  theme(axis.text.x = element_text(angle=70, size = 6, hjust = 1), legend.position = "bottom") +
  labs(x = "Category", y = "Proportion", linetype = "data set") +
  scale_y_continuous(labels = scales::percent) +
  geom_hline(yintercept = 0.5, alpha = 0.35) 
```

Overall, race demographics are consistent between our survey and post-stratification data sets. The proportion of Hispanic respondents in our survey and post-stratification data is comparable, at 12.7% and 13.5%, respectively. While there is a higher proportion of Hispanic respondents in our post-stratification data set, the difference is less than one percentage point. The proportions of white respondents in our survey (69.4%) and post-stratification (69.5%) data sets are nearly identical. @fig-stateprops illustrates the distribution of voters across the 50 states, including Washington D.C. in both our survey and post-stratification data sets. The steps and code for producing @fig-stateprops were obtained from @citeMatthew.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-stateprops
#| fig-cap: Survey and Post-Stratification Data Proportion of Voters by State

# steps obtained from matthewwankiewicz github repo US_election_forecast

survey_analysis_data$state <- tolower(survey_analysis_data$state)
state <- survey_analysis_data %>%
  group_by(state) %>%
  summarise(n = n()) %>%
  mutate(pct = n/sum(n), type = "survey",
         group = "state") %>%
   rename(level = state)

state_poststrat <- poststrat_analysis_data %>%
  group_by(state) %>%
  summarise(n = n()) %>%
  mutate(pct = n/sum(n), type = "post-stratification",
         group = "state") %>%
   rename(level = state)

voters_by_state <- rbind(state, state_poststrat)

voters_by_state |> ggplot(aes(as.factor(level), pct, group=as.factor(type), linetype = as.factor(type))) +
  geom_line() + facet_grid(~group, scales = "free") +
  theme(axis.text.x = element_text(angle=70, size = 6, hjust = 1)) +
  labs(x = "State", y = "% of Voters", linetype = "Data Set")+
  scale_y_continuous(labels = scales::percent)
```

We can see that there are some slight differences in the proportions of respondents in the survey and post-stratification data from Alabama, California, Florida, Kansas, Maryland, Massachussets, Michigan, New York, Nebraska, and Virginia. None of the voters in our survey data are from Vermont, but 0.198% of the voters in our post-stratification data are. Therefore, the results from applying our model to our post-stratification data will include a prediction for Vermont. @fig-stateprops shows that there is a comparable distribution of voters across U.S. states in the survey and post-stratification data. However, the survey data over represents Florida, Michigan, New York, and Virginia.

# Model {#sec-model}

We performed multi-level regression with post-stratification (MRP) to predict support for Joe Biden in the upcoming election. To perform MRP, we fit a logistic regression model to predict support for Biden using sex, age, race, race_hispanic, highest level of education, state, and urban as predictors on our survey data set [@americaspoliticalpulse], and post-stratified it using ACS census data [@citeIPUMS].

## Model set-up

We built our Bayesian logistic regression model using the `stan_glm()` function of the `rstanarm` package [@rstanarm]. The model that we use is:

```{=tex}
\begin{align} 
vote\_biden_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
\small
\mbox{logit}(\pi_i) &= \beta_0 + \beta_1 \text{sex}_i + \beta_2  \text{age\_bracket}_i + \beta_3 \text{race}_i +\beta_4 \text{race\_hispanic}_i+\beta_5 \text{education\_level}_i + \beta_6 \text{urban}_i \\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
\beta_4 &\sim \mbox{Normal}(0, 2.5) \\
\beta_5 &\sim \mbox{Normal}(0, 2.5) \\
\beta_6 &\sim \mbox{Normal}(0, 2.5)
\end{align}
```
where binary variable $vote\_biden_i$ is equal to 1 if the respondent's preferred 2024 presidential candidate is Joe Biden, or 0 if the respondent's preferred 2024 presidential candidate is Donald Trump. We run our model in R [@citeR] using `stan_glm`, with the default priors from `rstanarm` [@citerstanarm]. We then apply our model to our post-stratification data [@citeIPUMS] to predict the popular vote and electoral college results of the 2024 U.S. Presidential election.

Our model uses logistic regression, therefore one of its weaknesses is the fact that it predicts a binary outcome and does not consider the possibility that some American adults might vote for a third-party candidate or abstain from voting. Tradeoffs and concerns about our model are discussed in more detail in @sec-weaknesses-limitations.

## Model Justification

We expect to see a positive relationship between support for President Biden and non-white race, college or post-graduate education, and living in an urban area. This expectation is based on the fact that Trump has spent much of his political career spewing anti-immigrant rhetoric and America's urban/rural ideological divide. In 2016, Trump campaigned on the promise of building a wall at the U.S.-Mexico border, and during his presidency, Trump labeled Mexicans as rapists and drug smugglers, referred to the coronavirus as the "Kung-flu" and the "China virus." @citeGhitza found that 39% of Biden's supporters in the 2020 election were voters of color, compared to only 15% for Trump. Black and Asian-American voters turned out at an increased rate, heavily in favor of President Biden in 2020 [@citeGhitza]. Scala (2016) found that even after social, demographic, and economic factors including education, age, and race were controlled for, the urban-rural divide was still statistically significant in estimating vote patterns in U.S. presidential elections. The 2020 election was no exception, as Trump had strong support in rural areas and Biden dominated in major cities [@citeGhitza]. In @sec-survey, we found that respondents from rural areas were more likely to support Donald Trump, while respondents from urban areas were more likely to support Joe Biden.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# use this model to predict the following:
# overall support for biden
# support by state for biden
# electoral college map

# load the model with readRDS
presidential_vote_model <-
  readRDS(file = here::here("models/presidential_vote_model.rds"))
```

# Results {#sec-results}

## Popular Vote Prediction

We got our popular vote prediction by taking the weighted average of vote_biden across all U.S. states and converting it into a percent. The average is weighted according to the number of census respondents in our post-stratification data set from each state. @tbl-popvotepost-popularvote summarizes our mean, lower, and upper estimates for the popular vote percentage and number of electoral college votes that Joe Biden will receive on November 5th. Based on the results of our multi-level regression with post-stratification (MRP), which involved applying the logistic regression outlined in @sec-model to our post-stratification data set, we predict that Joe Biden and Donald Trump will receive 56.59% and 43.41%, respectively, of the popular vote. That means that our logistic regression model classified 56.59% of the people included in our subset of the ACS 2022 [@citeIPUMS] as Biden voters based on sex, age bracket, race, highest level of education, and whether they live in an urban or rural area. @tbl-popvotepost-popularvote shows our lower, mean, and upper estimate for the percent of the popular vote that Joe Biden will receive. Conversely, the lower estimate for Joe Biden corresponds to the upper estimate for Donald Trump, and the upper estimate for Joe Biden corresponds to the lower estimate for Donald Trump.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| include: true
#| label: tbl-popvotepost-popularvote
#| tbl-cap: "2024 U.S. election result estimates for Joe Biden based on post-stratification"
#| tbl-subcap: ["Popular Vote", "Electoral College"]

props <- poststrat_analysis_data %>%
  group_by(state, races, race_hispanic, age_bracket, sex, education_level, urban) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  group_by(state) %>%
  mutate(prop = n / sum(n))

num_per_state <- poststrat_analysis_data %>%
  group_by(state) %>%
  summarise(total_respondents = n())

biden_support_by_state <-
  presidential_vote_model |>
  add_epred_draws(newdata = props) |>
  rename(support_biden_predict = .epred) |>
  mutate(support_biden_predict_prop = support_biden_predict * prop) |>
  ungroup() |>
  summarise(support_biden_predict = sum(support_biden_predict_prop),
            .by = c(state, .draw)) |>
  summarise(
    mean = mean(support_biden_predict),
    lower = quantile(support_biden_predict, 0.025),
    upper = quantile(support_biden_predict, 0.975),
    .by = state
  )

merged_states <- merge(num_per_state, biden_support_by_state, by="state")

merged_states$mean_weighted_sum <- merged_states$total_respondents * merged_states$mean
merged_states$lower_weighted_sum <- merged_states$total_respondents * merged_states$lower
merged_states$upper_weighted_sum <- merged_states$total_respondents * merged_states$upper

biden_mean <- sum(merged_states$mean_weighted_sum) / sum(merged_states$total_respondents)
biden_lower <- sum(merged_states$lower_weighted_sum) / sum(merged_states$total_respondents)
biden_upper <- sum(merged_states$upper_weighted_sum) / sum(merged_states$total_respondents)

biden_popvote <- tibble(
  "Estimate:" = c("Lower Estimate", "Mean Estimate", "Upper Estimate"),
  "Biden %" = c(round(biden_lower * 100, 2), 
                round(biden_mean * 100, 2), 
                round(biden_upper * 100, 2)),
  "Trump %" = c(round((1 - biden_lower) * 100, 2), 
                round((1 - biden_mean) * 100, 2), 
                round((1 - biden_upper) * 100, 2))
)

# Set row names
rownames(biden_popvote) <- NULL

# electoral college
# had chatgpt make me an electoral college csv
electoral_college <- read_csv("electoral_college_votes.csv", show_col_types = FALSE)

electoral_college$electoral_votes <- as.numeric(electoral_college$electoral_votes)

merged_ec <- merge(biden_support_by_state, electoral_college, by = "state")

biden_states_mean <- merged_ec |>
  filter(mean >= 0.5)

trump_states_mean <- merged_ec |>
  filter(mean < 0.5)

biden_ec_mean <- sum(biden_states_mean$electoral_votes)
trump_ec_mean <- sum(trump_states_mean$electoral_votes)

biden_states_lower <- merged_ec |>
  filter(lower >= 0.5)
trump_states_lower <- merged_ec |>
  filter(lower < 0.5)

biden_ec_lower <- sum(biden_states_lower$electoral_votes)

biden_states_upper <- merged_ec |>
  filter(upper >= 0.5)

biden_ec_upper <- sum(biden_states_upper$electoral_votes)

biden_ec_poststrat <- tibble(
  "Electoral College Estimate:" = c("Lower Estimate", "Mean Estimate", "Upper Estimate"),
  "Biden" = c(biden_ec_lower, biden_ec_mean, biden_ec_upper),
  "Trump" = c(538-biden_ec_lower, 538-biden_ec_mean, 538-biden_ec_upper)
)

# Set row names
rownames(biden_ec_poststrat) <- NULL

# popular vote table
kable(biden_popvote, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = F)

# electoral college table
kable(biden_ec_poststrat, align = "c") %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

## Electoral College Prediction

The estimates for the support for Biden in each state are shown in @fig-estimatedpropsupport. We followed and modified the code provided in Chapter 16 of "Telling Stories with Data" [@tswd] to produce @fig-estimatedpropsupport. The red or blue line for the predicted proportion of Biden votes in each state represents the 97.5% prediction interval. The mean estimate for the percentage of voters in each state supporting Biden is denoted by the blue or red dot at the center of each line. The line is blue if the mean prediction is over 50%, indicating that the state's electoral college votes would go to Joe Biden, and red if it is below 50%, indicating that the state's elecotral college votes would go to Donald Trump. This interval crosses the 50% line for numerous states, which means that based on our model, they could be toss-ups. We can see that the prediction interval is widest for Hawaii, which ranges from 44.5% to 66.9% with the mean prediction of 56.1%, and Alaska, which ranges from 41.4% to 63.6%, with a mean prediction of 53.2%

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-estimatedpropsupport
#| fig-cap: "Estimated proportion of each state voting for Biden in 2024 (Post-Stratification)"

library(scales)

# we used the instructions provided in chapter 16 of telling stories with data to 
# produce and modifythis visualization

biden_support_by_state |>
  ggplot(aes(y = mean, x = fct_reorder(state, mean))) +
  geom_point(color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0, color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  labs(
    x = "State",
    y = "Estimated proportion support for Biden"
  ) +
  theme_classic() +
  scale_y_continuous(labels = scales::percent)+  # Setting y-axis to display percentages
  coord_flip() +
  theme(axis.text.y = element_text(size = 6)) 
```

@fig-estimatedpropsupport2 illustrates the same information as @fig-estimatedpropsupport, except it includes gray dots that represent the proportion of voters in each state that support Biden from the survey data. We followed and modified the code provided in Chapter 16 of "Telling Stories with Data" [@tswd] to produce @fig-estimatedpropsupport2.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-estimatedpropsupport2
#| fig-cap: "Estimated proportion of each state voting for Biden in 2024 Post-Stratification vs Survey Data"

survey_analysis_data$state <- tolower(survey_analysis_data$state)
biden_support_by_state |>
  ggplot(aes(y = mean, x = fct_reorder(state, mean))) +
  geom_point(color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0, color = ifelse(biden_support_by_state$mean > 0.5, "blue", "red")) +
  geom_point(
    data = survey_analysis_data |>
      summarise(n = n(),
                .by = c(state, vote_biden)) |>
      mutate(prop = n / sum(n),
             .by = state) |>
      filter(vote_biden == 1),
    aes(y = prop, x = state),
    color="#A9A9A9"
    ) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  labs(
    x = "State",
    y = "Estimated proportion support for Biden"
  ) +
  theme_classic() +
  scale_y_continuous(labels = scales::percent)+ # Setting y-axis to display percentages
  coord_flip() +
  theme(axis.text.y = element_text(size = 5)) 
```

We can see that there is a lot of variation between the post-stratification and survey estimates for the proportion of voters in each state that will vote for President Biden in the upcoming election. However, when we consider only the post-stratification estimates and add up the electoral college votes for the states in which the mean estimated support for Biden is over 50% (shown in blue on @fig-estimatedpropsupport and @fig-estimatedpropsupport2), we can predict that Joe Biden will defeat Donald Trump in the electoral college 471 to 67. @tbl-popvotepost-popularvote shows the electoral college predictions based on the lower end and higher end of these intervals. At the lower end, we can predict that Joe Biden will defeat Donald Trump in the electoral college 363 to 175, and at the upper end, we can predict that Joe Biden will defeat Donald Trump in the electoral college 538 to 0.

@fig-electoralmappoststrat illustrates the same information as @fig-estimatedpropsupport, but on an electoral map. We used the steps and code provided by @citeMatthew and the `statebins` package [@statebins] to produce @fig-electoralmappoststrat. This map shows the mean prediction for the proportion of voters in each state prefer Joe Biden over Donald Trump. We see that California (CA), Maryland (MD), and New Jersey (NJ) are dark blue, indicating strong support for Joe Biden in those states. Wyoming (WY), South Dakota (SD), North Dakota (ND), and Montana (MT) are red, which indicates support for Donald Trump. In addition to this, Texas and Florida are blue in @fig-electoralmappoststrat.The Democrats have not won Florida since former President Barack Obama won when he was re-elected in 2012. Donald Trump carried Florida in both 2016 and 2020, so flipping it in 2024 would be huge for President Biden. The Republican presidential nominee has carried the state of Texas in every presidential election since 1976 [@citeCNN], so flipping it in 2024 would give Biden an excellent chance of winning re-election.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-electoralmappoststrat
#| fig-cap: "Electoral map based on post-stratification data"

poststrat_electoral_map <- biden_support_by_state %>%
  mutate(prop = mean)

# this code comes from matthewwankiewicz 
# US_election_forecast/paper/forecasting_us_election.rmd

poststrat_electoral_map %>%
  mutate(statename = str_to_title(state)) %>%
  ggplot(aes(fill = prop, state = statename)) +
  geom_statebins() +
  scale_fill_gradient2(low = "#d12531", high = "#244999",
                       mid = "white", midpoint = mean(survey_electoral_map$prop)) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  labs(fill = "Proportion Voting \nfor Biden")
```

# Discussion {#sec-discussion}

## How we produced our popular vote and electoral college predictions {#sec-first-point}

## The popular vote does not necessarily define the electoral college

When discussing the results of our model, it is crucial to consider the quality of the survey data that was used to fit it. Despite projecting Joe Biden to win an impressing, but not dominant, 56.59% of the popular vote in the 2024 presidential election when applied to the post-stratification data, our model produced a mean estimate of 471 electoral college votes for Joe Biden. We know that winning a certain percentage of the popular vote does not translate into a fixed numer of electoral college votes. Instead, the candidate with the most votes in each state receives all of its electoral college votes, and in America's two-party system, the candidate whose electoral votes add up to 270 or more wins the presidency. So, in theory, it is possible to win 538 electoral votes with exactly 50.1% of the vote in each state. However, this scenario is extremely unlikely. It is also possible to win the popular vote, but still lose the electoral college and the presidency. In fact, Hillary Clinton won the popular vote but lost the electoral college to Donald Trump, 232 to 306, in 2016.

The shortcomings of our survey data set are highlighted in @fig-estimatedpropsupport2, where we can see that there is no visible pattern or relationship between predictions for the percent of voters that will support Biden in each state based on the survey and post-stratification data sets. The gray dots, which represent our prediction for the percent of voters who prefer Biden in each state based on our survey data set, are scattered all over the place. We used that same survey data set to fit the model that was applied to our post-stratification data set to produce the blue (where the mean prediction for Biden support was over 50%) and red (where the mean prediction for Biden support was below 50%) prediction intervals that show which candidate would win each state's electoral college votes. While state was not a predictor for our model, it is clear that in many cases, our survey and post-stratification data sets produced very different predictions for the percentage of votes that Biden will receive in many states.

The lower estimate for the number of electoral college votes that Biden will receive based on the results of our MRP in November is 363, while the upper estimate is 538, as shown in @tbl-popvotepost-popularvote. That is a difference of 175 seats, which gives us little confidence in our current model's ability to accurately predict the electoral college. In comparison, our prediction for the number of electoral college votes that President Biden will win in the 2024 election based on our survey data is 324, as shown in @tbl-surveyresults. 471 electoral college votes is not even relatively comparable to the 306 that he received in 2020, or the number of electoral college votes that presidents have received in recent elections [@citeCNN]. Therefore, there is reason for us to believe that the seemingly random and far-off predictions for the proportion of votes that Biden will receive in numerous states based on our survey data limits our model’s ability to accurately predict the electoral college.

## Weaknesses and Limitations {#sec-weaknesses-limitations}

Our model can only ever be as good as the data set that we used to build and fit it [@tswd]. The weaknesses of our model stem directly from the limited nature of the survey data set that we used, the America’s Political Pulse Survey Week 3 2024. As seen in @sec-survey, the electoral college predictions from Kansas (4 respondents), Wyoming (1 respondent), and Hawaii (7 respondents) are drastically different from the results seen in recent elections. There are very few respondents from smaller states with fewer electoral college votes, such as Kansas, Wyoming, Hawaii, New Hampshire, and Rhode Island in our survey data set. In addition to this, there are no respondents from Vermont in our survey data set. As mentioned, there were no respondents from Vermont and very few respondents from several states with few electoral college votes in our survey data set. As a result, we were not able to include state as a predictor in our model. Including state as a predictor would not have been applicable to respondents in our post-stratification census data from the state of Vermont at all.

In practice, including state and considering third-party candidates would make sense, but it was not worth the trade-offs for the America’s Political Pulse survey data set and our specific model. We initially did include state as a predictor, because we know that there are historical Democratic stronghold states, such as California, Illinois, Hawaii, Connecticut, Massachusetts, and Vermont, as well as historical Republican stronghold states such as Idaho, Utah, Kansas, Alabama, and Wyoming [@citeCNN]. However, this model resulted in warnings about the bulk effective sample size and the tail effective sample size being too low. This means that our model with state may have had unreliable posterior means, posterior medians, posterior variances, and tail quantiles [@citeStan]. This makes sense, as fitting a model that used state as a predictor on our specific survey data would have associated Kansas and Wyoming with support for President Biden and Hawaii with support for former President Trump. This would be reflective of the political preferences of fewer than five respondents from each of those states. It would fail to consider the historical Republican stronghold on Kansas and Wyoming and Democratic stronghold on Hawaii in presidential elections [@citeCNN].

Logistic regression can only be used for binary classification. This limits the scope of this paper, as it only allows us to consider support for Joe Biden and Donald Trump in the upcoming 2024 presidential election. Our model does not consider support for third-party candidates, write-in votes, or the possibility of not voting, which could be a disadvantage when predicting results in swing states or for close elections. In the 2016 election, Donald Trump defeated Hillary Clinton by 11,612 votes (0.3 percentage points) in Michigan, which is considered a swing state. Third-party candidates Gary Johnson (Libertarian) and Jill Stein (Green) each received over 173,000 votes (3.6% of the popular vote) and 50,700 votes (1.1% of the popular vote) in Michigan [@citePolitico2016]. This is just one example of at least one third-party candidate receiving more votes than the difference in votes between the Democratic and Republican nominees in a key battleground state. It highlights a major limitation of using logistic regression to predict vote preference.

## Next Steps

We can address our model's inability to consider more than two voting outcomes by fitting a SoftMax regression model. SoftMax regression is a generalization of logistic regression that can predict the class probabilities for more than two classics and perform multi-class classification. However, a SoftMax regression model is more powerful than a logistic regression model. It would severely overfit our current survey data set because it only has 1000 observations, including less than five from some smaller states. While a SoftMax regression model might make sense for predicting vote preference when there are more than two candidates, this model would have been far too powerful for our survey data set and not worth the trade-offs. The bulk effective sample size and tail effective sample size would have been too low, which would mean that our posterior means, posterior medians, posterior variances, and tail quantiles would be unreliable [@citeStan]. We can justify using a logistic regression model because the United States has a two-party system, meaning that Joe Biden (Democrat) and Donald Trump (Republican) are the only candidates who can realistically win the presidency on November 5^th^. Although there will be third-party candidates, we do not yet know who they will be, nor are they expected to win any electoral college votes.

With a survey data set that is as limited as the one we currently have, SoftMax regression is not a viable option. However, the possibility of third-party candidates receiving vote shares larger than the margin of victory of the Democratic or Republican nominee in key battleground states highlights the need to consider the fact that a small percentage of voters will cast a ballot for someone not named Donald Trump or Joe Biden on November 5th. Therefore, the first step towards extending and improving this report in the future would be to find a survey data set that has significantly more respondents, is somewhat reflective of voting trends and demographics in each state, and specifically asks respondents who they plan to vote for in the 2024 U.S. presidential election. Once we address the issue of having very few, or even no respondents from smaller states and no question about 2024 vote intention, we will be able to build a stronger and more accurate model to predict vote choice based on state, urban or rural area, demographics such as sex, age, and race, and education. A robust survey data set with enough respondents (possibly upwards of 10,000) will allow us to train a more powerful SoftMax regression model to perform multi-class classification and classify voters as Trump, Biden, third-party, or abstaining voters using similar predictors to the ones outlined in @sec-model as well as state. We believe that we can extend and improve our study by using SoftMax regression, but it is crucial for us to find a better survey data set first.

\newpage

\appendix

# Appendix {.unnumbered}

## Data Cleaning {#sec-datacleaning}

We created a binary variable, `vote_biden` to indicate whether each survey respondent's preferred 2024 presidential candidate is Joe Biden or Donald Trump. `vote_biden` is equal to 1, which indicates that the respondent's preferred presidential candidate is Joe Biden, if their party affiliation is Democratic or if they voted for Joe Biden in 2020. `vote_biden` is equal to 0 otherwise, indicating that a respondent's preferred presidential candidate is Donald Trump. Donald Trump is arguably the most polarizing figure in American politics, therefore we are considering independents who voted for Joe Biden in 2020 to prefer Joe Biden over Donald Trump again in 2024. However, vote_biden will never be equal to 1 if the respondent voted for Donald Trump in 2016 or 2020. In the context of this report and the vote_biden variable, voting for a third-party candidate, a write-in vote, or not voting are not considered possible voting outcomes.

We also created a new variable, `urban`. First, we added the `urban` variable to our survey data. It is equal to "urban" if `urbancity2` indicates that the respondent lives in a big city, a smaller city, or a suburb, and "rural" if the respondent lives in a small town or rural area. For our post-stratification data set, we also gave the `urban` variable two levels, "urban" and "rural" to align with our survey data set. If the `metro` variable indicated that a census respondent lives in a metropolitan area, whether inside or outside of the principal/central city, `urban` is equal to "urban," and we set it to rural for census respondents who do not live in a metropolitan area.

## Model Details

### Posterior Predictive Check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior.

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 1
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(presidential_vote_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(presidential_vote_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

### Markov Chain Monte Carlo Convergence

We check for signs that the Markov chain Monte Carlo (MCMC) algorithm ran into issues using an Rhat plot and a trace plot [@tswd].

@fig-stanareyouokay-1 is an Rhat plot. It shows everything is close than 1 and less than 1.05. This means that the coefficients all converge to the same distribution, and the model is able to predict vote_biden in the survey data.

@fig-stanareyouokay-2 shows the trace plots for our predictors. We can see that the lines bounce, but remain horizontal and have a lot of overlap with each other [@tswd]. This suggest that the distribution behaves as expected, and we do not need to re-run our model with fewer predictors or different priors [@tswd].

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Rhat plot", "Trace plot"]
#| layout-ncol: 1

plot(presidential_vote_model, "rhat") + theme_classic()

cat("\n")

plot(presidential_vote_model, "trace")

```

### Credibility Intervals

In @fig-modelplot, we show the 90% credibility intervals for the predictors of `vote_biden`.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-modelplot
#| fig-cap: "90% Credibility intervals for predictors of vote_biden"
modelplot(presidential_vote_model, conf_level = 0.9) +
  labs(x = "90% credibility interval")
```

\newpage

# References